Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-01-20 spie/paper.tex
draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications in fields such as architecture, conservation, and
simulation. Applying accurate textures to these models is important to
add context to digital visualizations, and allows for image-based
analysis of environments.

In this paper, we perform data acquisition with a backpack-mounted
system carried by an ambulatory human \cite{liu2010indoor}. This
backpack system contains a series of cameras, laser scanners, and
inertial measurement units, used to solve the simultaneous
localization and mapping (SLAM) problem, which produces backpack (and
therefore camera) poses over time as well as a point cloud
representing the surrounding environment \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure}. For environment geometry, we
work with models obtained by fitting low-resolution planar surfaces to
these point clouds \cite{sanchez2012point}.

@@

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
our geometry and explains a simple approach towards selecting images
for texturing. Section \ref{sec:existingApproaches} covers
sophisticated existing approaches to image stitching, and their
performance on our datasets. Section \ref{sec:2dAlignment} contains
our efficient approach towards 2D image alignment, followed by Section
\ref{sec:imageCompositing}, which describes our two methods of
compositing images. Section \ref{sec:resultsAndConclusions} contains
results and conclusions.

@@

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}.  We are provided with a set of
$M$ images to texture a target plane. Each image has a camera matrix
$P_i$ for $i=1..M$, which translates a 3D point in the world
coordinate system to a 2D point or pixel in image $i$'s coordinates. A
camera matrix $P_i$ is composed of the camera's intrinsic parameters,
containing focal length and image center, as well as extrinsic
parameters which specify the rotation and translation of the camera's
position in 3D world coordinates at the time that image $i$ is
taken. These extrinsic parameters are determined by the backpack
hardware and localization algorithms \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure} and are substantially noisy.

Because our backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the final texture, in order to minimize any visual discontinuities or
seams that would suggest that the plane's texture is not composed of a
single continuous image.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.5in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each plane through the
      camera matrics $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.125in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

@@

\section{Existing Approaches to Image Alignment}
Stitching together multiple images to produce a larger, seamless image
is a commonly performed task, with many successful approaches over the
past years. Generally, parts of images are matched to each other,
usually through direct comparisons or feature detection and
matching. Images are then transformed to maximize matches, often by
calculating homographies between pairs of images, or by iteratively
adjusting camera poses in 1 to 6 degrees of freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar data, and rotational
differences, as covered in X. Feature matching however, works best
when multiple unique visual references exist in the environment that
can be detected in multiple images. In contrast, our indoor
environments have a high prevalence of bare surfaces, as well as
repeating textures, such as similar windows, doors, and wall-mounted
decorations, that cause difficulty in disambiguating features. This
lack of reliable reference points often results in errors when
matching images together.

@@

In this section, we describe our method of efficient and robust image
alignment. Though state-of-the-art techniques towards image stitching
work in 3D, we elect to perform image alignments entirely in
2D. Besides the benifits of reduced complexity, we have found 2D
alignment to work well for two main reasons. First, because we have
such a large number of images to choose from, we have found that the
compositing approaches in Section \ref{sec:imageCompositing} is able
to select a subset of images that align well, given only 2D
alignments. Furthermore, the nature of our input data is such that
localization error chiefly occurs in two dimensions, in the same plane
as the surface being projected onto. Recall that our data acquisition
system is backpack-mounted, with cameras facing to the sides of the
operator. The operator makes efforts to walk parallel to walls, and
the localization and model-generation algorithms that provide our
input conform the operator's path to be straight and parallel to
detected surfaces \cite{kua2012loopclosure, sanchez2012point}. Since
the operator walks in as upright a position as possible, errors in
roll, yaw, and the operator's distance from parallel surfaces are
minimal. Thus, our highest errors stem from uncertainty in the
operator's pitch, equivalent to rotation around the camera axes, as
well as location along surfaces. These equate to 2D rotation and
translation in a surface's plane.

Our 2D alignment procedure consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then rotated and shifted to
match geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a linear least squares problem in 2D to maximize
matches.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms. Experience
and intuition show that walls in indoor environments often contain
linear features that are either horizontal or vertical, corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. This can be further generalized by instead orienting lines
to be parallel with a surface's boundaries, which makes it applicable
to non-rectangular surfaces, such as ceilings, floors, or slanted
walls.

@@

Furthermore, at this point in time, image occlusions have not been
accounted for. As a result, some image projections contain texture
that should project to an adjacent surface, sometimes with a visible
linear boundary where the two surfaces meet. If this linear boundary
is detected by Hough transform, the image is rotated and shifted in 2D
such that the visual boundary between two surfaces in an image
projection matches the physical boundary in our digital model. An
example of such an adjustment is in Figure \ref{fig:geometryAlignment}

To perform the rotation and alignment, we use the RANSAC
\cite{fischler1981random} framework. RANSAC allows us to determine an
optimal rotation and translation to match up two groups of line
segments, while accounting for outliers. The parameters we specify
were chosen after some tuning on our datasets, and are likely to vary
based on input data. First, a rotation angle is computed by gathering
all pairs of image line segments and geometry line segments with less
than 20 degrees between them. We then run the RANSAC algorithm by
sampling 4 of these angles at a time, and using their average as the
fitting function. Angles over 1 degree from this average are
considered outliers. Once a 2D rotation angle is determined, it is
applied to the image, and image line segments with length below a
certain threshold are discarded. Now, all pairs of image line segments
and geometry line segments with less than 0.1 degree difference in
orientation and less than 250 mm distance at their furthest points are
gathered. RANSAC is then similarly employed to calculate a shift that maximizes matches among these pairs of lines.

In an ideal case, homography might be applied here, but poop
In our datasets we have success sides for walls, floors and ceilings we have found often difficulty matching, so we often align only in X, and save this info later.
For floors and ceilings, we have more success, and since images smaller upwards, only one match usually, and so we save the slope for later.

optimal
horizontal and vertical translations independently, by sampling 4
pairs at a time and considering an outlier threshold of 10 mm. If a
translation is applied, we indicate that the image was aligned to
geometry in one or both dimensions, which will reduce further shifting
in Section \ref{sec:robustSIFTFeatureMatching}.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, nothing occurs. Where there is
a mixture of both, the rectangular region is subdivided into four, and
the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
parts of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, RANSAC
\cite{fischler1981random} is again used to estimate the optimal
$d^x_{i,j}$ and $d^y_{i,j}$ distances between two images $i$ and
$j$. For this application, the RANSAC fitting function finds the
average distance between selected matches in a pair of images, and the
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel threshold, so that
SIFT matches are labeled as outliers if their horizontal or vertical
distances are not within 10 pixels of the average distance computed by
the fitting function.

We now use the RANSAC-calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. There are a total of $M^{2}$ possible pairs of
images, though we only generate distances between images with SIFT
matches. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]

The $N \times M$ matrix $A$ is constructed with one row for each pair
of images with measured distances produced by the SIFT matching
stage. A row in the matrix has a $-1$ and $1$ in the columns
corresponding to the two images in the pair. For example, the matrix
below indicates a SIFT-based distance between images 1 and 2, images 1
and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0\\
\end{pmatrix}
\]
If only relative distances between images are included, the absolute
location of the images can not be calculated, and the matrix is rank
deficient. From Section \ref{sec:geometryAlignment}, a number of
images are anchored to geometry points, in one or both dimensions, and
thus their locations can be used to fix the rest in place. In case no
anchor images exist, we simply arbitrarily pick an image, as we have
no other reference points to work with. In the above matrix, the first
image is set to be such an anchor, simply by placing a $1$ in its
column.

The $N$-dimensional observation vector $\vec{\gamma}$ is constructed
using the SIFT-based distances generated in the RANSAC matching
stage. Elements in the observation vector corresponding to anchor
images are simply their locations as determined by the original noisy
localization. Thus a $\vec{\gamma}$ corresponding to the above matrix
can be written as:

\[
\vec{\gamma}^T =
\begin{pmatrix}
  d_{1,2}, &d_{1,3}, &d_{2,3}, &\hdots &d_{N-2,N-1}, &d_{N-1,N}, &x_1
\end{pmatrix}
\]

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. This solution however does not make use of our noisy camera
poses, and will fail when no SIFT matches are found between one
segment of the plane and another. To account for this, we add rows to
the $A$ matrix and observations to the $\vec{\gamma}$ vector
corresponding to the original noisy distances. We then add weighting
to our least squares problem where the SIFT distances and anchor
values are given a high weight e.g. 1, while the noisy distances are
given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that
section results in Figure \ref{fig:compareAll}(b), which has far fewer
discontinuities, though errors due to lighting differences and
repeating features are still visible.

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images, as
seen in Figures \ref{fig:compareAll}(e) and
\ref{fig:compareAll}(f). This shortest path approach approach directly
reduces the cost of each image boundary, while the tile caching method
uses a scoring function that only approximates this
effect. Furthermore, this approach guarantees the best selection of
images to minimize seams, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. This shortest path approach is
also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images.

@@

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(g)
and \ref{fig:compareAll}(h) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits much better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(h) clearly has the best visual quality among the
textures in Figure \ref{fig:compareAll}.

@@

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images, a
2.8GHz consumer-grade laptop takes approximately a minute to pick 36
candidate images, followed by another minute to perform both image
alignment and the shortest path texturing method, though over 75\% of
that time is spent calculating SIFT matches within the SiftGPU
framework. While not real-time, the process is capable of generating
quick updates after changes in various parameters or modifications to
input data, and if integrated directly into a modeling system, could
provide live visualization as data is collected.

In this paper, we have developed an approach to texture mapping models
with noisy camera localization data. We are able to refine image
locations based on geometry references and feature matching, and
robustly handle outliers. Using the tile-based mapping approach, we
can texture both simple rectangular walls as well as complex floor and
ceiling geometry. We also implemented a shortest path texturing method
that produces seamless textures on planes where multiple head-on
images are available. Both of these approaches are highly modular, and
easily tunable for different environments and acquisition systems.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications in fields such as architecture, conservation, and
simulation. Applying accurate textures to these models is important to
add context to digital visualizations, and allows for image-based
analysis of environments. In this paper, we perform data acquisition
with a backpack-mounted system carried by an ambulatory human
\cite{liu2010indoor}. This backpack system contains a series of
cameras, laser scanners, and inertial measurement units, used to solve
the simultaneous localization and mapping (SLAM) problem, which
produces backpack (and therefore camera) poses over time as well as a
point cloud representing the surrounding environment
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure}. For
environment geometry, we work with models obtained by fitting
low-resolution planar surfaces to these point clouds
\cite{sanchez2012point}.

@@

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
our geometry and explains a simple approach towards selecting images
for texturing. Section \ref{sec:existingApproaches} covers existing
approaches to image stitching, and their performance on our
datasets. Section \ref{sec:2dAlignment} contains our approach towards
efficient 2D image alignment, followed by Section
\ref{sec:imageCompositing}, which describes our two methods of
compositing images. Section \ref{sec:resultsAndConclusions} contains
results and conclusions.

@@

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}.  We are provided with a set of
$M$ images to texture a target plane, where each image has a camera
matrix $P_i$ for $i=1..M$, which translates a 3D point in the world
coordinate system to a 2D point or pixel in image $i$'s coordinates. A
camera matrix $P_i$ is composed of the camera's intrinsic parameters,
containing focal length and image center, as well as extrinsic
parameters which specify the rotation and translation of the camera's
position in 3D world coordinates at the time that image $i$ is
taken. These extrinsic parameters are determined by the backpack
hardware and localization algorithms \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure} and are substantially noisy.

Because our backpack system takes photos at a rate of 5 Hz, hundreds
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the final texture, in order to minimize any visual discontinuities or
seams that would suggest that the plane's texture is not composed of a
single continuous image.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.5in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each surface through the
      camera matrics $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.125in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

@@

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger, seamless image
is a commonly performed task, with many successful approaches over the
past years. Generally, parts of images are matched to each other,
usually through direct comparisons or feature detection and
matching. Images are then transformed to maximize matches, often by
calculating homographies between pairs of images, or by iteratively
adjusting camera poses in 1 to 6 degrees of freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar data, and rotational
differences \cite{szeliski2006image}. Feature matching however, works best
when multiple unique visual references exist in the environment that
can be detected in multiple images. In contrast, our indoor
environments have a high prevalence of bare surfaces, as well as
repeating textures, such as similar windows, doors, and wall-mounted
decorations, that cause difficulty in disambiguating features. This
lack of reliable reference points often results in errors when
matching images together.

@@

In this section, we describe our method of efficient and robust image
alignment. Though state-of-the-art techniques towards image stitching
work in 3D, we elect to perform image alignments entirely in
2D. Besides the benifits of reduced complexity, we have found 2D
alignment to work well for two main reasons. First, because we have
such a large number of images to choose from, the compositing
approaches in Section \ref{sec:imageCompositing} are able to select a
subset of images that align well, given only 2D alignments. Second,
the nature of our input data is such that localization error chiefly
occurs in two dimensions, in the same plane as the surface being
projected onto. Our data acquisition system contains cameras facing to
the sides of the operator. The operator makes efforts to walk parallel
to walls, and the localization and model-generation algorithms that
provide our input conform the operator's path to be straight and
parallel to detected surfaces \cite{kua2012loopclosure,
  sanchez2012point}. Since the operator walks in as upright a position
as possible, errors in roll, yaw, and the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

Our 2D alignment procedure consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then rotated and shifted to
match geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a weighted linear least squares problem in 2D to
maximize matches.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms, as shown in
Figure \ref{fig:geometryAlignment}(a). Experience and intuition show
that walls in indoor environments often contain linear features that
are either horizontal or vertical, corresponding to doors, windows,
posters, etc. Thus, when texturing walls, we rotate images in 2D such
that dominant lines are made to be horizontal or vertical. This can be
further generalized by instead orienting lines to be parallel with a
surface's boundaries, which makes it applicable to non-rectangular
surfaces, such as ceilings, floors, or slanted walls.

@@

Furthermore, at this point in time, image occlusions have not been
accounted for. As a result, some image projections contain texture
that should project to an adjacent surface, sometimes with a visible
linear boundary where the two surfaces meet. If this linear boundary
can be detected by Hough transform, the image is rotated and shifted
in 2D such that the visual boundary between two surfaces in an image
projection matches the physical boundary in our digital model. An
example of such an adjustment is in Figure
\ref{fig:geometryAlignment}(b) and \ref{fig:geometryAlignment}(c).

To perform rotation, we use a RANSAC \cite{fischler1981random}
framework to determine an optimal rotation angle between lines in our
image and lines in our geometry. RANSAC allows us to ignore erroneous
lines in our image, such as the slanted lines corresponding to the
ladder in Figure \ref{fig:geometryAlignment}(a). Once this rotation is
discovered and applied, we gather all pairs of image line segments and
geometry line segments with less than $0.1^\circ$ difference in
orientation and less than 250 mm distance at their furthest points. If
2 or more pairs exist, we select the 2 that contain the longest
noncollinear image-based line segments, and perform the corresponding
fixed translation to match them. If we only have 1 pair, we perform
the minimal shift to match the lines of that pair, and make note of
their orientation, for usage in Section
\ref{robustSIFTFeatureMatching}.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, the recursion stops. Where
there is a mixture of both, the rectangular region is subdivided into
four, and the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
parts of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, RANSAC
\cite{fischler1981random} is again used to estimate the optimal
$d^x_{i,j}$ and $d^y_{i,j}$ distances between two images $i$ and
$j$. We use a 10 pixel threshold for RANSAC, so that SIFT matches are
labeled as outliers if their distance is not within 10 pixels of the
sampled average distance.

We now use the $d^x_{i,j}$ and $d^y_{i,j}$ distances between each pair
of images to refine their positions using weighted linear least
squares. The variables we wish to solve for are the $x$ and $y$
positions of our images, while our equations are the SIFT-based
distances between pairs of images, images fixed to geometry with 0 or
1 degrees of freedom, and the original camera poses. Ignoring
weighting for now, an example setup for solving
$\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $
with 3 images is shown below.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  0 & -m & 0 & 0 & 1 & 0\\
\end{pmatrix}
\]

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & y_1, & y_2, & y_3
\end{pmatrix}
\]

\[
\vec{\gamma}^T =
\begin{pmatrix}
  dx_{1,2}, &dy_{1,2}, &dx_{2,3}, &dy_{2,3}, &t1_x, &t1_y, & -mx_2 +
  ty_2
\end{pmatrix}
\]

In this scenario, a SIFT-based distance of $dx_{1,2}$, $dy_{1,2}$ was
calculated between images 1 and 2. This corresponds to the first and
second of $A$, while the third and fourth row of $A$ represent the
same for images 2 and 3. Rows 5 and 6 of $A$ correspond to a fixed
location of $tx_1$, $ty_1$ for image 1, while Row 7 corresponds to a
constraint to a line of slope $m$, with current location $tx_2$,
$ty_2$, both results of the geometry alignment procedure in Section
\ref{sec:geometryAlignment}. If we do not have enough SIFT matches, or
lack images matched to geometry, our matrix becomes rank-deficient and
our problem cannot be solved. As a result, we add rows for each image
corresponding to the original noisy image locations, and downweight
them heavily, e.g. with a weight of 0.01 versus 1.

Because our problem is linear, it is easily and quickly solved, and after
applying the resulting shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that
section results in Figure \ref{fig:compareAll}(b), which has far fewer
discontinuities, though errors due to lighting differences and
repeating features are still visible.

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images. This
shortest path approach approach directly reduces the cost of each
image boundary, while the tile caching method uses a scoring function
that only approximates this effect. Furthermore, this approach
guarantees the best selection of images to minimize seams, while the
sequential tile caching method may select images early on that turn
out to be poor choices once subsequent tiles have been processed. This
shortest path approach is also far less intensive in terms of memory
usage and runtime, both during texture generation and rendering, as it
does not require discretizing planes or images.

@@

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. The
remaining images in Figure \ref{fig:compareAll} highlight differences
between the two methods, showing that Figure \ref{fig:compareAll}(f)
clearly has the best visual quality and the best texturing approach
among the textures in Figure \ref{fig:compareAll}.

@@

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images, a
2.8GHz consumer-grade laptop takes approximately a minute to pick 36
candidate images, followed by another minute to perform both image
alignment and the shortest path texturing method, though over 75\% of
that time is spent calculating SIFT matches within the SiftGPU
framework. While not real-time, the process is capable of generating
quick updates after changes in various parameters or modifications to
input data, and if integrated directly into a modeling system, could
provide live visualization as data is collected. Our full models
consist of our input model file, textures that we generate, and a
mapping of image points to model vertices. The models shown in Figure
\ref{fig:results} are roughly 20 MB in size, and are visualized using
the OpenSceneGraph toolkit \cite{openscenegraph}, which allows for export to many common
model formats, as well as efficient visualization, even in web
browsers or mobile devices.

In this paper, we have developed an approach to texture mapping models
with noisy camera localization data. We are able to refine image
locations based on geometry references and feature matching, and
robustly handle outliers. Using the tile-based mapping approach, we
can texture both simple rectangular walls as well as complex floor and
ceiling geometry. We also implemented a shortest path texturing method
that produces seamless textures on planes where multiple head-on
images are available. Both of these approaches are highly modular, and
easily tunable for similar systems across multiple environments.
